---
title: "Math 189 - Homework 1"
author: "Cathleen"
date: "4/19/2020"
output:
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 1


## 1. Summarize numerically the two distributions of birthweight for babies born to women who smoked during their pregnancy and for babies born to women who did not smoke during their pregnancy.

First, we will take a look at the statistics of birthweight looks and see a summary of the features.
```{r}
data <- read.table("babies.txt", header=TRUE)
summary(data$bwt)
``` 
```{r}
smoker.ind <- which(data['smoke'] == 1)
nonsmoker.ind <- which(data['smoke'] == 0)
data.smoker <- data[smoker.ind,]
data.nonsmoker <- data[nonsmoker.ind,]
```
#### Mean and Median birthweight of Nonsmoker mothers
```{r}
mean(data.nonsmoker$bwt)
median(data.nonsmoker$bwt)
```
#### Mean and Median birthweight of Smoker mothers
```{r}
mean(data.smoker$bwt)
median(data.smoker$bwt)
```
#### Summary Stats Nonsmoker vs. Smoker
```{r}
summary(data.nonsmoker$bwt)
summary(data.smoker$bwt)
```
#### Difference of Means and Medians between Smoker and Nonsmoker mothers
```{r}
mean(data.nonsmoker$bwt) - mean(data.smoker$bwt)
median(data.nonsmoker$bwt) - median(data.smoker$bwt)
```


## 2. Use graphical methods to compare the two distributions of birth weight.

#### Smoker vs Nonsmoker histogram distributions of birthweights.
Birthweight of smoking mothers is in red and nonsmoking mothers is in grey.
```{r}
library(ggplot2)
ggplot(data,aes(bwt)) + 
    geom_histogram(data=data.smoker, fill='red', alpha=.4) +
    geom_histogram(data=data.nonsmoker,', alpha=0.4)
```

#### Smoker and Nonsmoker density graph of birthweights.
Birthweight of nonsmoking mothers is in grey/black and smoking mothers is in red.

```{r}
ggplot(data, aes(bwt))+
  geom_density(data = data.smoker, color="red", fill="red", alpha=0.3) + 
  geom_vline(aes(xintercept=mean(data.smoker$bwt)),color="red", linetype="dashed", size=1) +
  geom_density(data = data.nonsmoker, color="black", fill="darkgrey", alpha=0.5) +
  geom_vline(aes(xintercept=mean(data.nonsmoker$bwt)),color="black", linetype="dashed", size=1)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
